Discovering governing reactions from concentration data

### 2018-01-22 docs/manuscript.tex
manuscript update

	When presented with a time series consisting of $T$ observations, the data can be represented as a matrix
	\begin{align}
	\textbf{X} = \begin{pmatrix}
		x_1(t_1) & x_2(t_1) & \cdots & x_S(t_1) \\
		x_1(t_2) & x_2(t_2) & \cdots & x_S(t_2) \\
		\vdots   & \vdots   & \ddots & \vdots   \\
		x_1(t_T) & x_2(t_T) & \cdots & x_S(t_T)
	\end{pmatrix} \in \mathbb{R}^{T\times S}.
	\end{align}
	Given this matrix, one can propose a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions with corresponding reaction functions
	\begin{align}
		\theta_r(\textbf{X}) = \begin{pmatrix}
		\textbf{y}_r(\textbf{X}_1)^T \\ \vdots \\ \textbf{y}_r(\textbf{X}_T)^T
		\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
	\label{method:the-reactions}\end{align}
	where $\textbf{X}_i$ denotes the $i$-th row in $X$. Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$. Following the approach of SINDy, the goal is now to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
	\end{pmatrix}^T$, so that
	\begin{align}
	\dot{\textbf{X}} = \Theta(\textbf{X})\Xi = \sum_{r=1}^{R}\theta_r(\textbf{X})\xi_r.
	\end{align}
	In particular, the system is linear in the coefficients $\Xi$, which makes sparse regression tools such as LASSO \cite{Tibshirani1996, Hastie2009} applicable. To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that
	\begin{align}
		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\|\Xi\|_1 \right) \quad \text{subject to }\Xi \geq 0,
	\label{method:minimizationproblem}\end{align}
	where $\|\cdot\|_F$ denotes the Frobenius norm and $\alpha\geq 0$ is a sparsity inducing hyperparameter. For $\alpha=0$ the minimization problem reduces to constrained least-squares. In order to solve (\ref{method:minimizationproblem}) we apply the sequential least-squares minimizer SLSQP, originally described in \cite{Kraft1988}, by the software package SciPy \cite{SciPy}. As initial guess for the regularized problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.

@@

	\begin{figure}
		\begin{center}
			\includegraphics[width=.5\textwidth]{./figures_tex/aux_linear_approximation.eps}
		\end{center}
		\caption{This figure depicts a possible output yielded by an application of the Gillespie method and the performed piecewise linear approximation in order to obtain a better behaving derivative.}
		\label{examples:linear-approximation}
	\end{figure}

>>>>

	When presented with a time series consisting of $T$ observations, the data can be represented as a matrix
	\begin{align}
	\textbf{X} = \begin{pmatrix}
		x_1(t_1) & x_2(t_1) & \cdots & x_S(t_1) \\
		x_1(t_2) & x_2(t_2) & \cdots & x_S(t_2) \\
		\vdots   & \vdots   & \ddots & \vdots   \\
		x_1(t_T) & x_2(t_T) & \cdots & x_S(t_T)
	\end{pmatrix} \in \mathbb{R}^{T\times S}.
	\end{align}
	Given this matrix, one can propose a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions with corresponding reaction functions
	\begin{align}
		\theta_r(\textbf{X}) = \begin{pmatrix}
		\textbf{y}_r(\textbf{X}_1)^T \\ \vdots \\ \textbf{y}_r(\textbf{X}_T)^T
		\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
	\label{method:the-reactions}\end{align}
	where $\textbf{X}_i$ denotes the $i$-th row in $X$. Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$. Following the approach of SINDy, the goal is to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
	\end{pmatrix}^T$, so that
	\begin{align}
	\dot{\textbf{X}} = \Theta(\textbf{X})\Xi = \sum_{r=1}^{R}\theta_r(\textbf{X})\xi_r.
	\end{align}
	In particular, the system is linear in the coefficients $\Xi$, which makes potentially sparse regression tools such as elastic net regularization \cite{Zou2005} applicable. To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that
	\begin{align}
		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right) \quad \text{subject to }\Xi \geq 0,
	\label{method:minimizationproblem}\end{align}
	where $\|\cdot\|_F$ denotes the Frobenius norm, $\lambda\in[0,1]$ a hyperparameter that interpolates linearly between LASSO \cite{Tibshirani1996, Hastie2009} and Ridge \cite{Hoerl1} methods, and $\alpha\geq 0$ is a hyperparameter that, depending on $\lambda$, can induce sparsity and give preference to smaller solutions in the $L_1$ or $L_2$ sense.

	For $\alpha=0$ the minimization problem reduces to constrained least-squares. In order to solve (\ref{method:minimizationproblem}) we apply the sequential least-squares minimizer SLSQP, originally described in \cite{Kraft1988}, by the software package SciPy \cite{SciPy}. As initial guess for the regularized problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.

@@

	%\begin{figure}
	%	\begin{center}
	%		\includegraphics[width=.5\textwidth]{./figures_tex/aux_linear_approximation.eps}
	%	\end{center}
	%	\caption{This figure depicts a possible output yielded by an application of the Gillespie method and the performed piecewise linear approximation in order to obtain a better behaving derivative.}
	%	\label{examples:linear-approximation}
	%\end{figure}

	(\textcolor{red}{this has changed, rewrite!})
